Sensor Clusterization in D-optimal Design in Infinite Dimensional Linear Bayesian Inverse Problems

### 2020-10-02 new_paper.tex
Actually tested the algorithm for finding AAt with unit norm columns. It worksgit add zeros.py  Flawlesslygit add zeros.py !

\subsection{D-Optimal Design in Infinite Dimensions}\label{subsec:D optimal design} 
The meaning of D-optimal design in infinite-dimensional Hilbert spaces
was investigated in \cite{AlexanderianGloorGhattas14}. Note that the
authors make assumptions that amount to $\Sigma=I$ (implied by
$\modcov = 0,\sigma^2=1$), but we choose not to take these
simplification here. This is because $\Sigma$ can determine ``how
much'' clusterization we see. As stated
\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
general covariance matrices. The conclusion is that in
infinite-dimensions, a D-optimal design is well-defined as maximizing
the expected KL divergence between posterior and prior (denoted $\tar$
below and throughout the paper) and is summarized, using our notation,
in the following theorem:
\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
  Let $\pr = \normal(0,\prcov)$ be a centered Gaussian prior
  on $\hil$   and let $\post = \normal(\postmean,\postcov)$ 
  the posterior measure on $\ban$ for the Bayesian linear
  inverse problem $\data =  \obs \fwd\param + \obs \eps' + \eps$ discussed
  above. Then 
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}

@@

First variation of $\tar (\obs)$ in the direction $V$ is:
\begin{align*}%%\label{eq:tar var}
  \begin{split}
    \delta \tar(\obs) V 
    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
    % 
    % 
    % 
    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by definition \eqref{eq:objective})}\\
    % 
    % 
    % 
    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
    \obs\fwd \prcov^{1/2} )^{-1}
    \frac{\der}{\der \tau} \Big |_{\tau=0}
    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
    % 
    % 
    % 
    &= \frac12 \ttr{\postcov \fwd^* (V^* \Sigma^{-1} \obs 
      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
      &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
      + \obs^* \Sigma^{-1} V ) \fwd}  \text{ (by \ref{lemma:aux calc})}\\
%%   \end{split}
%% \end{align} 
%% Taking adjoints and cyclic permutations we see that
%% \begin{align*}
%%   \tr \postcov \fwd^* V^* \Sigma^{-1} \obs \fwd 
%%   &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V \fwd,
%% \end{align*}
%% and
%% \begin{align*}
%%   \tr \postcov \fwd^* \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \fwd
%%   &= \tr \postcov \fwd^*  \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs \fwd.
%% \end{align*}
%% We use these in the first equality of the following derivation,
%% which continues the one in \eqref{eq:precursor}.
%% %
%% \begin{align} \label{eq:tar var}
%%   \begin{split}
      %% \delta \tar(\obs) V 
    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V 
    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
    %
    %
    % 
    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
    % 
    %
    %
    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
  \end{split}
\end{align*} 
Denote
\begin{align}\label{eq:tar grad}
  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
  \postcov \fwd^* \obs^*\Sigma^{-1},
\end{align}
the gradient of $\tar (\obs )$ and we now justify this definition.
Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
just $\tr{A} = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
basis vector), we see that
\begin{align*}
  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
  V_j(\nabla \tar(\obs)_j),
\end{align*}
with $V_j \in \ban^*$ and $\nabla \tar(\obs)_j \in \ban^{**}\subseteq
\ban, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in (\ban^{**})^m
\subseteq \ban^m$ is indeed the correct gradient and the notation
\eqref{eq:tar grad} is justified. Note that by the comment following
\eqref{eq:obs*}, we view $V$ (defined on the same space as $\obs$) as
a column vector. Then the gradient $\nabla \tar(\obs)$ should be
viewed as a row vector, as the product $V \nabla \tar$ is in $\R^{m
  \times m}$. This will prove important in section
\ref{subsec:necessary}.

@@

\subsection{The Eigenvalue Problem}\label{subsec:eigenproblem}
When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
first-order condition for optimality \eqref{eq:conditions} becomes
\begin{equation}\label{eq:eigenproblem}
  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi, 
\end{equation}
with $\Xi$ diagonal. This is an eigenvalue problem for the
self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Below,
we use \eqref{eq:postcov} in the second equality. If $\fwd$ is
invertible, the third equality is trivial. Otherwise, we use some
calculations delegated to lemma \ref{lemma:twice woodbury} in the
appendix:
\begin{align}\label{eq:mod conditions}
  \begin{split}
  \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \\
  %
  %
  %
  &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^* \\
  %
  %
  %
  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
  \end{split}
\end{align}  
This presents the optimal design problem as a problem on $\ban$ and
not on $\hil$ --- the term $(\fwd \prcov \fwd^*)^{-1}$ is the prior
precision in $\ban$ and $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
\obs$ is the posterior precision in $\ban$.

\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
$\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq \dots$ and its
eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma (proved
in the appendix \ref{lemma:sim diag}) will be useful:

@@

Thus, if we take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
\fwd^*)^{-1}$ in the above lemma, we conclude from the eigenproblem
\eqref{eq:mod conditions} that $\obs^*\obs$ has the same eigenvectors
as $\fwd \prcov \fwd^*$, namely $\{\ev_i\}_{i=1}^{\infty}$ (infinitely
many of them with zero eigenvalue, since $\obs^*\obs$ is
finite-rank). Denote the corresponding non-zero eigenvalues of
$\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
\geq k+1$.

\begin{align}\label{eq:pre manipulations}
  \begin{split}
    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
    \obs \fwd \prcov^{1/2}\right ) &=
    %
    \det \left (I + \sigma^{-2} \obs^*
    \obs \fwd \prcov\fwd^* \right ) \text{(Sylvester's determinant theorem)}\\
    %
    %
    %
    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}\\
  \end{split}
\end{align}

By definition of the objective \eqref{eq:objective} and using
\eqref{eq:pre manipulations}:
\begin{align}\label{eq:mod objective}
  \begin{split}
    \tar ( \obs ) &=\frac12 \log \det \left( I + \sigma^{-2}
    \prcov^{1/2} \fwd ^* \obs^* \obs \fwd \prcov^{1/2} \right )\\
    %
    %
    %
    %
    &=\frac12 \log \det \Big ( \left ( (\fwd \prcov \fwd^* )^{-1} +
    \sigma^{-2}\obs^* \obs \right ) \fwd \prcov \fwd^* \Big )\\
    %
    %
    %
    &=\frac12 \log \left (\prod_{i=1}^{\infty} ( \lambda_i^{-1} + \sigma^{-2} \eta_i ) \prod_{i=1}^{\infty} \lambda_i \right ) \\
    %
    %
    %
    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right ). 
  \end{split}
\end{align}
Therefore, an optimal design maximizes

\begin{equation}\label{eq:true target}
  \sum_{i=1}^{k}\log (\lambda_i^{-1} + \sigma^{-2} \eta_i),
\end{equation}

@@

      \addlegendentry{prior}
      \addlegendentry{$o_1$}
      \addlegendentry{$o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \caption{Posterior precision per eigenvector, after taking a
    measurement $\meas_i$. Each $\meas_i$ increases precision
    uniformly across the lowest precision eigenvectors. We see that
    $\meas_2$ and $\meas_3$ are repeated measurements, as well as
    $\meas_5$ and $\meas_6$.}
  \label{fig:clusterization}
\end{figure}
with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
appendix (Lemma \ref{lemma:free}) we can find $\meas_1,\dots,\meas_m$
such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
to the restriction that we do not change the trace of $\obs^*\obs$, so
that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the fastest
increase in the design criterion (equation \eqref{eq:true target}) is
gained by increasing $\lambda_i^{-1} + \sigma^{-2}\eta_i$ where it is
smallest. So, no weight should be given to (i.e. measurement taken of)
the $k$th eigenvector before $\lambda_i^{-1} + \sigma^{-2}\eta_i \geq
\lambda_k^{-1},\ \forall i \neq k$. Thus, we should choose
$\{\eta_i\}_{i=1}^k$ as follows. We would increase $\lambda_1^{-1} +
\sigma^{-2} \eta_1$ until it equals $\lambda_2^{-1}$. Then,
$\lambda_i^{-1} + \sigma^{-2} \eta_i,i=1,2$ are increased until
$\lambda_i^{-1} + \sigma^{-2} \eta_i = \lambda_3^{-1},i=1,2$ and so
forth.

The argument above has the important consequence that the eigenvalues
in \eqref{eq:eigenproblem} are all equal This makes perfect
sense. First, we know the Lagrange multipliers tell us how much we
gain by relaxing the constraints and there is no reason for any
measurement to give us more than any other (otherwise we would put
more weight on the better measurement, as can be understood from the
discussion in the following paragraph).%%  We are also used to symmetric
%% optimization problems having symmetric solutions \cite{Waterhouse83},
%% so symmetric and identical Lagrange multipliers definitely make sense
%% in this context.

\subsection{Sensor Clusterization}\label{subsec:clusterization}
The way clusterization can occur is best understood from figure
\ref{fig:clusterization}. We see that each measurement increases
precision (reduces uncertainty) only for the eigenvectors of lowest
precision. If the precision in these is lower than the precision for
the next eigenvector, a repeated measurement is optimal.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
\subsection{Overview}
In this section we show how clusterization is avoided when model
error does not vanish.

\subsection{Derivation}
We rely on corollary \ref{cor:same meas} below, which is proved in the
appendix. Denote $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
(\meas_1,\dots,\meas_{m-1})^t$.
\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}.
  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
  \begin{equation*}
    \tar(\obs) - \tar(\obsm) =
    \log \left ( 1 + \frac{\sigma^2
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ).
  \end{equation*}
\end{restatable*}
 In
contrast to the previous case of vanishing model error, we will see
that if $\modcov \neq 0$ clustering will not occur. From Corollary
\ref{cor:same meas} we can observe that

>>>>

\subsection{D-Optimal Design in Infinite Dimensions}\label{subsec:D optimal design} 
The meaning of D-optimal design in infinite-dimensional Hilbert spaces
was investigated in \cite{AlexanderianGloorGhattas14}. Note that the
authors make assumptions that amount to $\Sigma=I$ (implied by
$\modcov = 0,\sigma^2=1$), but we choose not to take these
simplification here. This is because $\Sigma$ can determine ``how
much'' clusterization we see. As stated
\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
general covariance matrices. The conclusion is that in
infinite-dimensions, a D-optimal design is well-defined as maximizing
the expected KL divergence between posterior and prior and is summarized, using our notation,
in the following theorem:
\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
  Let $\pr = \normal(0,\prcov)$ be a centered Gaussian prior
  on $\hil$   and let $\post = \normal(\postmean,\postcov)$ 
  the posterior measure on $\ban$ for the Bayesian linear
  inverse problem $\data =  \obs \fwd\param + \obs \eps' + \eps$ discussed
  above. Then 
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}
A D-optimal design seeks to maximize $\tar(\obs)$.

@@

First variation of $\tar (\obs)$ in the direction $V$ is:
\begin{align*}%%\label{eq:tar var}
  \begin{split}
    \delta \tar(\obs) V 
    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
    % 
    % 
    % 
    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by definition \eqref{eq:objective})}\\
    % 
    % 
    % 
    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
    \obs\fwd \prcov^{1/2} )^{-1}
    \frac{\der}{\der \tau} \Big |_{\tau=0}
    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
    % 
    % 
    % 
    &= \frac12 \ttr{\postcov \fwd^* (V^* \Sigma^{-1} \obs 
      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
      &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
      + \obs^* \Sigma^{-1} V ) \fwd}  \text{ (by \ref{lemma:aux calc})}\\
%%   \end{split}
%% \end{align} 
%% Taking adjoints and cyclic permutations we see that
%% \begin{align*}
%%   \tr \postcov \fwd^* V^* \Sigma^{-1} \obs \fwd 
%%   &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V \fwd,
%% \end{align*}
%% and
%% \begin{align*}
%%   \tr \postcov \fwd^* \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \fwd
%%   &= \tr \postcov \fwd^*  \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs \fwd.
%% \end{align*}
%% We use these in the first equality of the following derivation,
%% which continues the one in \eqref{eq:precursor}.
%% %
%% \begin{align} \label{eq:tar var}
%%   \begin{split}
      %% \delta \tar(\obs) V 
    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V 
    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
    %
    %
    % 
    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
    % 
    %
    %
    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
  \end{split}
\end{align*} 
Denote
\begin{align}\label{eq:tar grad}
  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
  \postcov \fwd^* \obs^*\Sigma^{-1},
\end{align}
the gradient of $\tar (\obs )$ and we now justify this definition.
Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
basis vector), we see that
\begin{align*}
  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
  V_j(\nabla \tar(\obs)_j),
\end{align*}
with $V_j \in \ban^*$ and $\nabla \tar(\obs)_j \in \ban^{**}\subseteq
\ban, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in (\ban^{**})^m
\subseteq \ban^m$ is indeed the correct gradient and the notation
\eqref{eq:tar grad} is justified. Note that by the comment following
\eqref{eq:obs*}, we view $V$ (defined on the same space as $\obs$) as
a column vector. Then the gradient $\nabla \tar(\obs)$ should be
viewed as a row vector, as the product $V \nabla \tar$ is in $\R^{m
  \times m}$. This will prove important in section
\ref{subsec:necessary}.

@@

\subsection{The Eigenvalue Problem}\label{subsec:eigenproblem}
When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
first-order condition for optimality \eqref{eq:conditions} becomes
\begin{equation}\label{eq:eigenproblem}
  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi, 
\end{equation}
with $\Xi$ diagonal. This is an eigenvalue problem for the
self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Below,
we use the expression for the posterior covariance \eqref{eq:postcov}
in the second equality. If $\fwd$ is invertible, the third equality is
trivial. Otherwise, we use some calculations delegated to lemma
\ref{lemma:twice woodbury} in the appendix:
\begin{align}\label{eq:mod conditions}
  \begin{split}
  \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \\
  %
  %
  %
  &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^* \\
  %
  %
  %
  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
  \end{split}
\end{align}  
This presents the optimal design problem as a problem on $\ban$ and
not on $\hil$ --- the term $(\fwd \prcov \fwd^*)^{-1}$ is the prior
precision in $\ban$ and $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
\obs$ is the posterior precision in $\ban$.

\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
$\fwd \prcov \fwd^*$ by $\{\lambda_i\}_{i=1}^{\infty}$ and its
eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma (proved
in the appendix \ref{lemma:sim diag}) is crucial for the rest of the
analysis:

@@

Thus, if we take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
\fwd^*)^{-1}$ in the above lemma, we conclude from the eigenproblem
\eqref{eq:mod conditions} that $\obs^*\obs$ has the same eigenvectors
as $\fwd \prcov \fwd^*$, namely $\{\ev_i\}_{i=1}^{\infty}$ (all but
finitely many of them with zero eigenvalue, since $\obs^*\obs$ is
finite-rank). Denote the corresponding non-zero eigenvalues of
$\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
\geq k+1$. Using Sylvester's determinant theorem we get:

\begin{align}\label{eq:pre manipulations}
  \begin{split}
    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
    \obs \fwd \prcov^{1/2}\right ) &=
    %
    \det \left (I + \sigma^{-2} \obs^*
    \obs \fwd \prcov\fwd^* \right ) \\
    %
    %
    %
    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}\\
  \end{split}
\end{align}

By definition of the objective \eqref{eq:objective} and using
\eqref{eq:pre manipulations}:
\begin{align}\label{eq:mod objective}
  \begin{split}
    \tar ( \obs ) &=\frac12 \log \det \left( I + \sigma^{-2}
    \prcov^{1/2} \fwd ^* \obs^* \obs \fwd \prcov^{1/2} \right )\\
    %
    %
    %
    %
    &=\frac12 \log \det \Big ( \left ( (\fwd \prcov \fwd^* )^{-1} +
    \sigma^{-2}\obs^* \obs \right ) \fwd \prcov \fwd^* \Big )\\
    %
    %
    %
    &=\frac12 \log \left (\prod_{i=1}^{\infty} ( \lambda_i^{-1} + \sigma^{-2} \eta_i ) \prod_{i=1}^{\infty} \lambda_i \right ) \\
    %
    %
    %
    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )\\
    %
    %
    %
    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), \text{ up to an additive constant}. 
  \end{split}
\end{align}

Therefore, an optimal design maximizes

\begin{equation}\label{eq:true target}
  \sum_{i=1}^{k}\log (\sigma^2 \lambda_i^{-1} + \eta_i),
\end{equation}

\pgfplotstableread{
  Label     prior  optimal  sub-optimal 
  1         0.2    1.3           1.4
  2         0.8    0.7           0.3
  3         2.2    0             0.2
  4         3.5    0             0.1
}\optimalvsnot

\begin{figure}
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\optimalvsnot}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};

      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
      \addlegendentry{$\eta_i$ from an optimal design}

    \end{axis}
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\optimalvsnot}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};

      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
      \addlegendentry{$\eta_i$ from a sub-optimal desgin}

    \end{axis}
  \end{tikzpicture}
  \caption{Characterizing an optimal design, compared to a
    sub-optimal design. Both designs are allowed two measurements,
    which means the blue area has accumulated height of $2$ in both
    plots. An optimal design (left) increases precision (reduces
    uncertainty) where it is lowest. A sub-optimal design (right) does
    not. In this example, a measurement is taken partly in the direction of the
    fourth eigenvector, where precision is highest.}
  \label{fig:optimal vs not}
\end{figure}

@@

      \addlegendentry{prior}
      \addlegendentry{$o_1$}
      \addlegendentry{$o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \caption{Posterior precision per eigenvector, after taking a
    measurement $\meas_i$. Each $\meas_i$ increases precision
    uniformly across the lowest precision eigenvectors. We see that
    $\meas_2$ and $\meas_3$ are repeated measurements, as well as
    $\meas_5$ and $\meas_6$.}
  \label{fig:clusterization}
\end{figure}

with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
appendix (Lemma \ref{lemma:free}) we can find $\meas_1,\dots,\meas_m$
such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
to the restriction that we do not change the trace of $\obs^*\obs$, so
that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the fastest
increase in the design criterion (equation \eqref{eq:true target}) is
gained by increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is
smallest. So, no weight should be given to (i.e. measurement taken in
direction of) the $k$th eigenvector before $\sigma^2\lambda_i^{-1} +
\eta_i \geq \sigma^2\lambda_k^{-1},\ \forall i \neq k$. Thus, we
should choose $\{\eta_i\}_{i=1}^k$ as follows. First, order the
eigenvalues $\lambda_1 \geq \lambda_2 \geq \dots$. We should increase
$\sigma^2 \lambda_1^{-1} + \eta_1$ until it equals
$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} + \eta_i,i=1,2$ are
increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
\sigma^2\lambda_3^{-1},i=1,2$ and so forth.

The argument above has the important consequence (illustrated in
figure \ref{fig:optimal vs not}) that the eigenvalues in the
eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
perfect sense. First, we know the Lagrange multipliers tell us how
much we gain by relaxing the constraints and there is no reason for
any measurement to give us more than any other (otherwise we would put
more weight on the better measurement, as can be understood from the
discussion in the following paragraph). This result of the analysis
above is summarized in the theorem below.

\begin{theorem}[Characterizing D-optimal designs]\label{thm:char}
In a D-optimal design $\eta_i, i=1,\dots,k$ are chosen such that
$\sigma^2\lambda_i^{-1} + \eta_i < \sigma^2\lambda_k^{-1}$ and
$\sigma^2\lambda_i^{-1} + \eta_i$ are all equal for $1 \leq i \leq k$
for some $k$. For $i \geq k+1$, however, $\eta_i = 0$.
\end{theorem}

%% We are also used to symmetric optimization problems having
%% symmetric solutions \cite{Waterhouse83}, so symmetric and identical
%% Lagrange multipliers definitely make sense in this context.

\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
\subsection{Overview}
In this section we explain how sensor clusterization may arise. We
first consider a sequential optimal design, where measurements are
chosen sequentially. We then consider the case where all measurements
are chosen at once. Finally, we hypothesize how this may apply to a
real-life inverse problem.

\subsection{Sequential Design}
The way clusterization can occur is illustrated in figure
\ref{fig:clusterization} and understood using theorem
\ref{thm:char}. The first measurement $\meas_1$ is optimal, since
$\eta_i, i=1,2$ are chosen to increase and equalize
$\sigma^2\lambda_i^{-1} +\eta_i, i=1,2$, which are lowest. In this
case, $\meas_1$ has no component in the direction of $\ev_i, i\geq
3$. We see that each measurement increases precision (reduces
uncertainty) only for the eigenvectors of lowest precision. Consider a
sequential optimization scenario. We first seek a measurement to
maximize our our utility, given the prior. From , the posterior
covariance does not depend on data observed, so we know it after
deciding on the first measurement.

the best
first measuremen, thIf the precision in these is lower than the
precision for the next eigenvector, a repeated measurement is optimal.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
\subsection{Overview}
In this section we show the effect model error has on the
clusterization phenomenon. In contrast to the previous case of
vanishing model error, we will see that if $\modcov \neq 0$ clustering
will not occur.

\subsection{Derivation}
We rely on corollary \ref{cor:same meas} below, which is proved in the
appendix. Denote $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
(\meas_1,\dots,\meas_{m-1})^t$.
\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
  \begin{equation*}
    \tar(\obs) - \tar(\obsm) =
    \log \left ( 1 + \frac{\sigma^2
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ).
  \end{equation*}
\end{restatable*}
From Corollary \ref{cor:same meas} we can deduce that
